An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-01-08 Articles/Article_EnerEcon_2014/EconGrowth4.tex
Created EconGrowth4 and began making Martin's edits.

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}

@@

\journal{Energy Economics}

\begin{document}

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconModels)
require(EconData)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(nlmrt)
require(dplyr)
# Provides access to data melting. 
# See http://cran.r-project.org/web/packages/reshape2/reshape2.pdf
require(reshape2) 
# tikz allows use of LaTeX formatting and font in graphs. 
# Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions 
# on installing tikzDevice.
# require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=TRUE)
require(knitr)

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

<<constants, include=FALSE, eval=TRUE>>=
spaghettiGraphLevel <- 1.0 # Spaghetti graph bands will encompass a certain percentage of the data.
@

<<data>>=
# Load historical data
historicalData <- Calvin # From the EconData package

# The working directory is set to the directory that contains this file.
# To reach the post-processed data, we need to move 
# up two levels and into the data_postprocessed directory.
# So, in each of the loadPostProcessedData calls below, we use
# dir = "../../data_postprocessed".
pp_path <- file.path("..", "..", "data_postprocessed")

# Load coefficients from all models in one large data frame
coeffs_df <- loadPostProcessedData(Source = "Calvin", kind = "coeffs", dir = pp_path)
# Load fitted data from all models in one large data frame
fitted_df <- loadPostProcessedData(Source = "Calvin", kind = "fitted", dir = pp_path)
# Load the original models.
oModels <- loadPostProcessedData(Source = "Calvin", kind = "models", dir = pp_path)
@

@@

Figure~\ref{fig:Historical_Graph} presents historical data, 
including indexed values for 
output ($y$), 
capital stock ($k$), 
labor ($l$), and 
energy ($q$) 
for 1980--2011 (developed economies) and 1991--2011 (other economies). 
**** Remove word ``Calvin'' from top. --MKH ****
%
<<Historical_Graph, fig.pos="H", fig.align='center', fig.width=3, fig.height=7.5, fig.cap="Historical data: indexed GDP ($y$), capital stock ($k$), labor ($l$), and thermal energy ($q$).">>=
mdata <- melt(historicalData, 
              measure.vars = c("iGDP", "iK", "iL", "iQp"), 
              value.name = "value", 
              variable.name = "variable") 
historicalPlot(data=mdata, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ Source, 
               line_types = c(1,5,2,3)) +
  xlab("") +
  ylab("Indexed (1980=1 or 1991=1)")
@

@@

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
using data for each economy (see Sections~\ref{sec:Economy_Types} and~\ref{sec:Sources_of_Data}).
We obtained parameter estimates by applying the method of least-squares to 
log-transformed data%
  \footnote{
  As is typical in the literature, we employ the natural logarithm transform
  to allow linear fitting of exponential coefficients.
  }
using the \texttt{R} \citep{R} functions
\texttt{lm} (for Cobb-Douglas and LINEX models) 
and \texttt{cesEst} \citep{Henningsen:2011td} (for CES models).

@@

Figure~\ref{fig:CD_GDP_Spaghetti_Graph} compares predictions 
from the Cobb-Douglas models 
(without and with energy) to historical data.
Historical data are shown as a black line. 
The fit to historical data is shown as a white line.
The gray band encompasses \Sexpr{spaghettiGraphLevel * 100.0}\% of the fits to resampled data.

<<CD_lambda_theta_graph, fig.pos="H", fig.align='center', fig.width=2.8, fig.height=7.0, fig.cap="$\\lambda$ and $\\theta$ values for CD resample models.">>=
cdCoeffsData <- filter(coeffs_df, model=="cd", nestStr %in% c("iK+iL", "iK+iL+iQp"))
standardScatterPlot(data=cdCoeffsData, mapping=aes(x=scale, y=lambda), facet_formula=country~energy, alpha=0.05) +
    scale_x_continuous(limits=c(0.8, 1.2), breaks=c(0.8, 1.0, 1.2)) +
    scale_y_continuous(limits=c(-0.05, 0.15), breaks=c(0, 0.05, 0.1)) +    
    labs(x=expression(theta), y=expression(lambda))
@

\twocolumn{}
<<CD_Ternary_Scatter_Graph, fig.pos="H", fig.align='center', fig.width=2.9, fig.height=7.7, fig.cap="Cobb-Douglas resample model parameters.">>=
standardTriPlot(data=cdCoeffsData, facet_formula=country~energy)
@

<<CD_GDP_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=3, fig.height=8.2, fig.cap="Cobb-Douglas fitted models.">>=
cdFittedData <- filter(fitted_df, model=="cd", nestStr %in% c("iK+iL", "iK+iL+iQp"))
spaghettiPlot(data=cdFittedData, facet_formula=Country~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@
\onecolumn{}

@@

<<CES_lambda_gamma_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="$\\lambda$ and $\\gamma$ values for CES resample models.">>=
cesCoeffsData <- filter(coeffs_df, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
                                                                     "(iK + iL) + (iQp)", 
                                                                     "(iL + iQp) + (iK)", 
                                                                     "(iQp + iK) + (iL)"))
standardScatterPlot(data=cesCoeffsData, mapping=aes(gamma_coef, lambda), facet_formula=country~nestStrParen, alpha=0.05) +
    scale_x_continuous(breaks=c(0.9, 1.0, 1.1)) +
    scale_y_continuous(limits=c(-0.05, 0.15), breaks=c(0, 0.05, 0.1)) +    
    labs(x=expression(gamma), y=expression(lambda))
@

<<CES_sigma1_delta1_graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="$\\sigma_1$ and $\\delta_1$ values for CES resample models.">>=
standardScatterPlot(data=cesCoeffsData, mapping=aes(delta_1, sigma_1), facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta[1]), y=expression(sigma[1]))
@

<<CES_sigma_delta_graph, fig.pos="H", fig.align='center', fig.width=4.5, fig.height=7, fig.cap="$\\sigma$ and $\\delta$ values for CES resample models.">>=
standardScatterPlot(data=filter(cesCoeffsData, nestStrParen != "(iK + iL) + ()"),
                    mapping=aes(delta, sigma), facet_formula=country~nestStrParen, alpha=0.05) +
    coord_trans(y="sigma") +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    scale_y_continuous(breaks=c(0, 0.5, 1.0, 2.0)) +
    labs(x=expression(delta), y=expression(sigma))
@

@@

<<CES_ternary_graph, fig.pos="H", fig.align='center', fig.width=5.5, fig.height=7.5, fig.cap="$\\alpha$, $\\beta$, and $\\gamma$ values for CES resample models.">>=
standardTriPlot(data=cesCoeffsData, facet_formula=country~nestStrParen)
@

Figure~\ref{fig:CES_Spaghetti_Graph} compares CES model predictions 
to historical data for all nestings. 
As with the Cobb-Douglas results, 
historical data are shown as a black line, 
the fit to historical data is shown as a white line, and 
the gray band encompasses \Sexpr{spaghettiGraphLevel * 100.0}\% of the fits to resampled data.

<<CES_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=8.5, fig.cap="CES fitted models.">>=
cesFittedData <- filter(fitted_df, model=="ces", nestStrParen %in% c("(iK + iL) + ()", 
                                                                     "(iK + iL) + (iQp)", 
                                                                     "(iL + iQp) + (iK)", 
                                                                     "(iQp + iK) + (iL)"))
spaghettiPlot(data=cesFittedData, facet_formula=Country~nestStrParen, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@

@@

Figure~\ref{fig:LINEX_Spaghetti_Graph} compares LINEX model predictions to historical data.
Again, historical data are shown as a black line, 
the fit to historical data is shown as a white line and 
the gray band encompasses \Sexpr{spaghettiGraphLevel * 100.0}\% of the fits to resampled data.

\twocolumn{}
<<LINEXq_Resample_Graph, fig.pos="H", fig.align='center', fig.width=2, fig.height=7, fig.cap="LINEX model (with energy) resampling results.">>=
linexCoeffsData <- filter(coeffs_df, model=="linex", energy=="iQp")
standardScatterPlot(data=linexCoeffsData, aes(a_0, c_t), facet_formula=country~energy, alpha=0.05) +
    labs(x=expression(a[0]), y=expression(c[t])) +
     scale_x_continuous(limits=c(0, 2), breaks=c(0, 1.0, 2.0)) +
     scale_y_continuous(limits=c(0, 10), breaks=c(0, 5, 10))
@

<<LINEX_Spaghetti_Graph, fig.pos="H", fig.align='center', fig.width=2, fig.height=7.1, fig.cap="LINEX fitted models.">>=
linexFittedData <- filter(fitted_df, model=="linex", energy=="iQp")
spaghettiPlot(data=linexFittedData, facet_formula=Country~energy, level=spaghettiGraphLevel) +
    labs(y="Indexed GDP (1980=1 or 1991=1)")
@
\onecolumn{}

<<abg_trajectories, eval=FALSE>>=
# This code can be run to test the conclusion about the LINEX model meeting constraints.
# Need to put oModels in the environment much earlier in this file.
# This chunk of R code has eval=FALSE, because it is not a part of the paper.
abgplotdata <- fortify(oModels$Calvin$IR$linex$`iK+iL+iQp`)
triPlot(data=abgplotdata, grid_lines=5, x=alpha, y=beta, z=gamma, geom=geom_line, aes_string="colour = iYear")
@

>>>>

%% This is file `elsarticle-template-2-harv.tex',
%%
%% Copyright 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle  Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: elsarticle-template-2-harv.tex 155 2009-10-08 05:35:05Z rishi $
%% $URL: http://lenova.river-valley.com/svn/elsbst/trunk/elsarticle-template-2-harv.tex $
%%
\documentclass[preprint,authoryear,12pt]{elsarticle}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

\journal{Energy Economics}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: replacing previous import by 'scales::rescale' when loading 'EconModels'}}\end{kframe}
\end{knitrout}

@@

Figure~\ref{fig:Historical_Graph} presents historical data, 
including indexed values for 
output ($y$), 
capital stock ($k$), 
labor ($l$), and 
energy ($q$) 
for 1980--2011 (developed economies) and 1991--2011 (other economies). 
**** Remove word ``Calvin'' from top. --MKH ****
%

@@

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
using data for each economy (see Sections~\ref{sec:Economy_Types} and~\ref{sec:Sources_of_Data}).
We obtained parameter estimates by applying the method of least-squares to 
log-transformed data using the \texttt{R} \citep{R} functions
\texttt{lm} (for Cobb-Douglas and LINEX models) 
and \texttt{cesEst} \citep{Henningsen:2011td} (for CES models).%
  \footnote{
  As is typical in the literature, we employ the natural logarithm transform
  to allow linear fitting of exponential coefficients.
  }

@@

Figure~\ref{fig:CD_GDP_Spaghetti_Graph} compares predictions 
from the Cobb-Douglas models 
(without and with energy) to historical data.
Historical data are shown as a black line. 
The fit to historical data is shown as a white line.
The gray band encompasses ??\% of the fits to resampled data.

\twocolumn{}

\onecolumn{}

@@

@@

Figure~\ref{fig:CES_Spaghetti_Graph} compares CES model predictions 
to historical data for all nestings. 
As with the Cobb-Douglas results, 
historical data are shown as a black line, 
the fit to historical data is shown as a white line, and 
the gray band encompasses ??\% of the fits to resampled data.

@@

Figure~\ref{fig:LINEX_Spaghetti_Graph} compares LINEX model predictions to historical data.
Again, historical data are shown as a black line, 
the fit to historical data is shown as a white line and 
the gray band encompasses ??\% of the fits to resampled data.

\twocolumn{}

\onecolumn{}
